---
title: "Analysis of hunting effects on spatial structure of trees with different dispersers"
subtitle: "Data preparation"
author: "Robert Bagchi"
date: "28 December, 2016"
output: html_document
---

```{r global_options, include=FALSE}
options(width=80, digits=3)
```


## 1. Preliminaries

### 1.1 Loading and installing libraries

If you need to install the `ReplicatedPointPatterns` package, you can execute the following code. Note, you need the `devtools` package installed. 

```{r package_install, eval=FALSE}
devtools::install_github('robertbagchi/ReplicatedPointPatterns')
```

We can then load all the necessary libraries.

```{r loadlibs, warning=FALSE, message=FALSE}
rm(list=ls())
# data organisation
library(abind)

library(reshape)
library(tidyr)
library(dplyr)
library(broom)

## spatial analysis
library(spatstat)
library(ReplicatedPointPatterns)

## plotting
library(ggplot2)
library(cowplot)
library(grid)
library(gtable)
```


### 1.2 Define some useful functions for this analysis


* Function that standardises a vector (standardise)
* Function to convert K functions to L functions (K2L)
* Function that takes the output from the model and puts it in a convenient format for plotting (kfunclme2plot)
* Function that plots the BLUEs against distance (plot.kfunctionlme)
* Function that plots the coefficient and confidence intervals from a bootstrap (effectplot.Kfunctionlme)
* Function to extract random effect BLUPs to allow plotting (extractModRanefs)
* Function to extract residuals and covariates for plotting (plotModResids)

```{r convenienceFuncs}
source("PeruConvenienceFunctions.R")
```

## 2. Read in and process data

### 2.1 Read in the data
```{r loaddata}
trees <- read.csv("../../data/mdd6plots_data_22jul2016.csv") ## tree locations
dispersal <-  read.csv("../../data/MDDspp_dispersal_syndromes_22jul2016.csv") ## dispersal data
sitedat <- read.csv("../../data/sitedata.csv") ## site information
```

```{r definitions}
#Define maximum range over which to consider spatial patterns
rmax <- 15
```

### 2.2 Data organisation

#### 2.1.1 Organise the dispersal data
```{r disporg}
# Put dispersal into 1/0 format by filling in the NAs with 0s
dispersal[, c('LV', 'SV', 'SB', 'Bat', 'TR', 'Wind', 'Expl', 'Unkwn', 'HSD', 'HID')] <-
  apply(dispersal[, c('LV', 'SV', 'SB', 'Bat', 'TR', 'Wind', 'Expl', 'Unkwn', 'HSD', 'HID')],
        2, function(x) ifelse(is.na(x), 0, x))
# combine wind and explosive dispersal into a single variable
dispersal$Abiotic <-  dispersal$Wind + dispersal$Expl
```

#### 2.2.2 Calculate hunting pressure
Construct hunting pressure variable as the negative sum of the standardised densities of large primates and large birds
```{r huntprescalc}
sitedat$hunted <- relevel(sitedat$hunted, 'intact')
sitedat$huntpres <- with(sitedat, (standardise(lg.primates) + standardise(lg.birds) + standardise(md.primates)))*(-1)
sitedat[order(sitedat$pname),] ## take a quick look
```

#### 2.2.3 Organise spatial pattern data

Need to deal with colocated points and remove stems that are 
a.  palms,  
b. unidentified (spcode = -999)
c. have no dispersal data
d. Juveniles > 5 cm dbh

We also remove from the bivariate analysis species without at least two individuals in each cohort (adults, juveniles and saplings) in the central ha of the site and  from the univariate analysis species which have <2 juveniles or saplings in the central ha of a site

```{r spporg}
# First deal with duplicated points (same site, coordinates, code and Type) -
# sometimes these have different tags, but for our purposes, 2 points at the
# same location cause problems, so we add a little bit of noise to them
# (less than the resolution of the original data)
dups <- duplicated(trees[,c('Site', 'E', 'N', 'code', 'Type')])
sum(dups)## a total of 102 such points out of 67000, so not many
trees$E[dups]  <-  trees$E[dups] + runif(sum(dups), -0.1, 0.1)
trees$N[dups]  <-  trees$N[dups] + runif(sum(dups), -0.1, 0.1)

trees.old <- trees ## make a copy for comparison later on
## Remove palms
trees <-trees[-c(grep('Astrocaryum', trees$Spp),
                 grep('Attalea', trees$Spp),
                 grep('Iriartea', trees$Spp),
                 grep("Chamaedorea", trees$Spp)),] ## removing palms

nrow(trees.old) - nrow(trees) ## lose 3727 palms


## remove unidentified stems
sum(trees$code==-999) ## removes 7154 unidentified stems
trees <-  trees[trees$code != -999,] 
nrow(trees) ## leaves us with 55870 stems

## now remove juveniles between 5-10 cm
## there are a few stems with missing dbh data - remove them too
trees <- trees[!(trees$diam > 5 & trees$Type=='J'),]
trees <- trees[!is.na(trees$Type),]
nrow(trees) ## lose 3008 stems. Seems relatively minor.

## Now get the abundance of each species x type x site combination
## and merge it into the dataset
abund <- aggregate(Tag ~ Spp + code + Type + Site, data=trees, length)
names(abund)[5] <- 'abund'

abund <- spread(abund, Type, abund, fill=0)
names(abund)[4:6] <- paste('N', names(abund[4:6]), sep='.')
abund$hasdisp <- abund$code %in% dispersal$code
## now working out if there is enough replication at each 
## species and site combination for 

abund$Quni <- abund$N.S > 1 & abund$N.J > 1 & abund$hasdisp
abund$Qbi <- abund$N.A > 1 & abund$N.J > 1 & abund$N.S > 1 & abund$hasdisp

## Some preliminary numbers of how many stems and species qualify
sum(abund$Quni); sum(abund$Qbi) ## 601 & 437 species
apply(abund[abund$Quni, c('N.J', 'N.S')], 2, sum) ## 17470 juvs and 9709 saps 
apply(abund[abund$Qbi, c('N.A', 'N.J', 'N.S')], 2, sum) ## 8820 ads 14093 juvs and  7942 saps 


trees <- merge(trees, abund, by=c('Spp', 'code', 'Site'), all =T)
trees$Spp <- droplevels(trees$Spp)
nlevels(trees$Spp)
summary(trees)
```


### 2.3 Calculate the K functions

Define the plot boundaries

```{r plotbounds}
## defining the window at CC12 which is not quite rectangular
cc12  <- owin(poly=list(x=c(225, 75, 75, 24.9, 25, 0, 0, 225),
                        y= c(200, 200, 125, 125, 75, 75, 0, 0)))

## defining a list of windows for each site
## Each comprises of a window for each size class.
winlist <- list(BM=list(A=owin(c(0, 200), c(0, 200)),
                        J=owin(c(50, 150), c(50, 150)),
                        S=owin(c(50, 150), c(50, 150))),
                CashuTr12=list(A=cc12,
                               J=owin(c(75, 175), c(50, 150)),
                               S=owin(c(75, 175), c(50, 150))),
                CashuTr3=list(A=owin(c(0, 200), c(0, 200)),
                              J=owin(c(55, 145), c(55, 145)),
                              S=owin(c(55, 145), c(55, 145))),
                LA=list(A=owin(c(0, 200), c(0, 200)),
                        J=owin(c(50, 150), c(50, 150)),
                        S=owin(c(50, 150), c(50, 150))),
                RA=list(A=owin(c(0, 200), c(0, 200)),
                        J=owin(c(50, 150), c(50, 150)),
                        S=owin(c(50, 150), c(50, 150))),
                TRC=list(A=owin(c(0, 200), c(0, 200)),
                         J=owin(c(50, 150), c(50, 150)),
                         S=owin(c(50, 150), c(50, 150))))

par(mfrow=c(2,3))
lapply(winlist, function(x) { plot(x$A, axes=T, asp=1)
  plot(x$J, add=T, border=2)}) ## looks ok
```

Split the data by site and pull out species lists

```{r sitesplit}
trees <- split(trees, f=trees$Site)

## Pull out some summary stats
sapply(trees, nrow)
sapply(trees, function(x) length(unique(x$Spp)))

## make a species list for each plot (based on the species that have sufficient
## stems in each size class 
## Univariate analysis
splists.uni <- lapply(trees, function(dat) unique(dat$code[dat$Quni]))

## bivariate analysis
splists.bi <- lapply(trees, function(dat) unique(dat$code[dat$Qbi]))
```

Turn the data from each species into a point pattern, split up by age class
A few adult points in CC12 are right on the boundary and this leads to the warnings below, which we deal with in the final function which removes species where this is a problem from the analysis.

```{r pointpatterns}
ppp.sps.uni <- mapply(function(spid, treedat, win){
  ppp.sp <- lapply(as.list(spid), function(id, treedat, win){
    treedat <- droplevels(treedat[treedat$Type !='A',])
    win <- win[c('J', 'S')]
    condat <- treedat[treedat$code==id,] ## subset data from species
    ppp.con <- mapply(function(Tx, Wx){ ## turn into ppp object
      pppx <- ppp(x=Tx$E, y=Tx$N, window=Wx)
      return(pppx)}, Tx=split(condat, f=condat$Type), Wx=win, SIMPLIFY=FALSE)
    ## note that species are split by size-class (Type) when input to the function
    ## now heterospecifics
    hetdat <- treedat[treedat$code!=id,] ## subset out all heterspecifics
    ## Note that this will include species
    ## not included in elsewhere in the analysis
    ppp.het <- mapply(function(Tx, Wx){
      Tx <- Tx[!duplicated(Tx[,c('E', 'N')]),] ## remove colocated stems
      pppx <- ppp(x=Tx$E, y=Tx$N, window=Wx)
      
      return(pppx)}, Tx=split(hetdat, f=hetdat$Type), Wx=win, SIMPLIFY=FALSE)
    names(ppp.het) <- paste(names(ppp.het), 'h', sep='') ## add 'h' suffix to het names
    ppp.sp <- c(ppp.con, ppp.het) ## combine cons and hets
    return(ppp.sp)},
    treedat=treedat, win=win)
  names(ppp.sp) <- spid
  return(ppp.sp)
}, spid=splists.uni, treedat=trees, win=winlist, SIMPLIFY=FALSE)

## Now for the bivariate case.
ppp.sps.bi <- mapply(function(spid, treedat, win){
  ppp.sp <- lapply(as.list(spid), function(id, treedat, win){
    condat <- treedat[treedat$code==id,] ## subset data from species
    ppp.con <- mapply(function(Tx, Wx){ ## turn into ppp object
      pppx <- ppp(x=Tx$E, y=Tx$N, window=Wx)
      return(pppx)}, Tx=split(condat, f=condat$Type), Wx=win, SIMPLIFY=FALSE)
    ## note that species are split by size-class (Type) when input to the function
    ## now heterospecifics
    hetdat <- treedat[treedat$code!=id,] ## subset out all heterspecifics
    ## Note that this will include species
    ## not included in elsewhere in the analysis
    ppp.het <- mapply(function(Tx, Wx){
      Tx <- Tx[!duplicated(Tx[,c('E', 'N')]),] ## remove colocated stems
      pppx <- ppp(x=Tx$E, y=Tx$N, window=Wx)
      
      return(pppx)}, Tx=split(hetdat, f=hetdat$Type), Wx=win, SIMPLIFY=FALSE)
    names(ppp.het) <- paste(names(ppp.het), 'h', sep='') ## add 'h' suffix to het names
    ppp.sp <- c(ppp.con, ppp.het) ## combine cons and hets
    return(ppp.sp)},
    treedat=treedat, win=win)
  names(ppp.sp) <- spid
  return(ppp.sp)
}, spid=splists.bi, treedat=trees, win=winlist, SIMPLIFY=FALSE)


## Remove species with adults only on the edge of the plot
ppp.sps.bi <-  
  lapply(ppp.sps.bi, function(site.dat){
    site.dat[sapply(site.dat, function(sp.dat){
      npoints(sp.dat$A[dilation(Window(sp.dat$J), 15)])}) > 0]})
```

## 3. K function calculations
We can now calculate the K functions for both the univariate and bivariate analyses. 

### 3.1 Univariate K functions

The code below constructs a hyperframe (see the spatstat package) which includes the covariates, point patterns and K functions in a single object. 
```{r univariate_kfuncs}
hyperdat.uni  <- mapply(function(dat, win, sitenms)
    {

        hyper <- hyperframe(
          site=sitenms,
          sp.id = rep(names(dat), 2), ## 2 times for S+J
          stage = factor(rep(c('S', 'J'), each=length(dat)), levels=c('S', 'J')),
          pppx=c(lapply(dat, function(x) x$S),
                 lapply(dat, function(x) x$J)),
          pppy=c(lapply(dat, function(x) x$Sh),
                 lapply(dat, function(x) x$Jh))
        )
        ## Add site data
        hyper <- cbind.hyperframe(hyper,
          sitedat[match(hyper$site, sitedat$site),
                  c('forest', 'pname', 'hunted', 'lg.primates', 'lg.birds', 'huntpres')])
        
        hyper$N1 <- sapply(hyper$pppx, npoints) # no. conspecifics
        hyper$N2 <- sapply(hyper$pppy, npoints) # no. heterospecifics
      
        # Add dispersal data
        hyper <- cbind.hyperframe(hyper,
                                  dispersal[match(hyper$sp.id, dispersal$code),
                                            c('Spp', 'LV', 'SV', 'SB', 'Bat',
                                              'TR', 'Wind', 'Expl',
                                              'Unkwn', 'Abiotic', 'HSD', 'HID')])
        ## Concpecific K function
        Kcon <- with(hyper, Kest(pppx, r=0:rmax, correction="border", 
                                    ratio=TRUE))
        ## Heterospecific K function
        ppp.s <- with(hyper, superimpose(i=pppx, j=pppy, W=pppy$window))
        Khet <- lapply(ppp.s, function(ppp.s)
          Kmulti(ppp.s, I=marks(ppp.s)=='i', J= marks(ppp.s)=='j',
                 r=0:rmax, correction="border", ratio=TRUE))
        
        ## Correct conspecific K function for inhomogeneities in
        ## overall tree density by subtracting Khet from Kcon
        hyper$K <-  mapply(function(conK, hetK){
          K <- conK
          K$border <- conK$border - hetK$border
          K$theo <- conK$theo - hetK$theo
          return(K)}, conK=Kcon, hetK=Khet, SIMPLIFY=FALSE)
      
        hyper$wts <-  lapply(hyper$pppx, function(pppx)
            kfunc.weights.calc(pppx=pppx, r=0:rmax,
                               correction='border', type='sqrtnxny_A'))
                            
        return(hyper)
    }, dat = ppp.sps.uni, win=winlist, sitenms=as.list(names(ppp.sps.uni)), 
    SIMPLIFY=FALSE)
```

We now do a final cut of species x site combinations with insufficient data. Even if there are enough points, too many individuals may be on the border and so are removed when border corrections are applied. We need at least one individual within the inner (minus-sampled) region. The easiest
way to do this is to find `NAs` in the K functions and remove them 
<!-- way to do this is run the models because the function has an inbuilt algorithm that discards species with fewer individuals than a given cut-off. -->

```{r selectspeciesuni, warnings=FALSE}
## First look for any NAs in the Kfunctions - which is a clue of insufficient data
rmlist <- lapply(hyperdat.uni, function(d)
  d$sp.id[sapply(d$K, function(dk) any(is.na(dk$border)))])

hyperdat.uni.sel <- mapply(function(dat, sprm){
  dat <- dat[!(dat$sp.id %in% sprm),]
  return(dat)}, dat=hyperdat.uni, sprm=rmlist, 
  SIMPLIFY=FALSE) 


## gives us the number of speciesxsite combinations and for each site.
sum(sapply(hyperdat.uni.sel, function(x) length(unique(x$Spp))))
sapply(hyperdat.uni.sel, function(x) length(unique(x$Spp)))

## make one hyperframe with all the data
hyperdat.uni.sel <- do.call('rbind', hyperdat.uni.sel)
## We can now remove unused levels
hyperdat.uni.sel$sp.id <- droplevels(hyperdat.uni.sel$sp.id)
hyperdat.uni.sel$Spp <- droplevels(hyperdat.uni.sel$Spp)
summary(hyperdat.uni.sel)
```


### 3.2 Bivariate K-functions

The bivariate calculations are similar to the univariate ones, except we are now interested in adult neighbours.

```{r bivariate_kfuncs}
hyperdat.bi  <- mapply(function(dat, sitenms)
{
  hyper <- hyperframe(
    site = sitenms,  
    sp.id = rep(names(dat), 2), ## 2 times for S+J 
    stage = factor(rep(c('S', 'J'), each=length(dat)), levels=c('S', 'J')),
    pppx=c(lapply(dat, function(x) x$S),
           lapply(dat, function(x) x$J)),
    pppy = rep(lapply(dat, function(x) x$A), 2),
    ppphet = rep(lapply(dat, function(x) x$Ah), 2)
  )
  
  hyper <- cbind.hyperframe(hyper,
                            sitedat[match(hyper$site, sitedat$site),
                                    c('forest', 'pname', 'hunted', 
                                      'lg.primates', 'lg.birds', 'huntpres')])
  
  hyper$N1 <- sapply(hyper$pppx, npoints)
  hyper$N2 <- sapply(hyper$pppy, npoints)
  
  hyper <- cbind.hyperframe(hyper,
                            dispersal[match(hyper$sp.id, dispersal$code),
                                      c('Spp', 'LV', 'SV', 'SB', 'Bat', 'TR',
                                        'Wind', 'Expl',
                                        'Unkwn', 'Abiotic','HSD', 'HID')])
  ppp.s <- with(hyper, superimpose(i=pppx, j=pppy, W=pppy$window))
  
  Kcon <- lapply(ppp.s, function(x)
    Kmulti(x, I=x$marks =='i', J=x$marks=='j',
           r=0:rmax, corr='border', ratio=TRUE))
  
  ppp.het <- with(hyper, superimpose(i=pppx, j=ppphet, W=ppphet$window))
  Khet <- lapply(ppp.het, function(x)
    Kmulti(x, I=x$marks =='i', J=x$marks=='j',
           r=0:rmax, corr='border', ratio=TRUE))
  hyper$K <- mapply(function(conK, hetK){
    K <- conK
    K$border <- conK$border - hetK$border
    K$theo <- conK$theo - hetK$theo
    return(K)}, conK=Kcon, hetK=Khet, SIMPLIFY=FALSE)
  
  hyper$wts <-  mapply(function(pppx, pppy)
    kfunc.weights.calc(pppx=pppx, pppy=pppy, r=0:rmax,
                       correction='border', type='sqrtnxny_A'),
    pppx=hyper$pppx, pppy=hyper$pppy, SIMPLIFY=FALSE)
  
  return(hyper)
}, dat=ppp.sps.bi, sitenms=as.list(names(ppp.sps.bi)),
SIMPLIFY=FALSE)

hyperdat.bi <- do.call('rbind', hyperdat.bi)
summary(hyperdat.bi)
```


Because we are using plus-sampling rather than minus-sampling for the bivariate analysis, we don't need to do a final cut of speices as we did in the univariaten case. 



Later analysis reveals that some species/site combinations are major outliers in the bivariate analysis. We could remove them in individual analyses, but that starts to get messy and could lead to errors. To deal with this problem once and for all, we remove them here
```{r removeoutliers}
## Removing those species now
hyperdat.bi.sel <- subset(hyperdat.bi, !((sp.id == 108 & pname=='BM') |
                                           (sp.id==314 & pname=='CC2') |
                                           (sp.id==38 & pname=='LA') |
                                           (sp.id==144 & pname=='LA') |
                                           (sp.id==226 & pname=="CC2") |
                                           sp.id == 712 & pname=="TRC")
)

## We can now remove unused levels
hyperdat.bi.sel$sp.id <- droplevels(hyperdat.bi.sel$sp.id)
hyperdat.bi.sel$Spp <- droplevels(hyperdat.bi.sel$Spp)

```

## 4. Finalise datasets
At this point it is useful to save the objects we have created and are required for analysis so we don't always have to repeat processing and to minimise processing on the cluster. We save our results as a data object.

```{r saveobjects}
save('dispersal', "sitedat", 'hyperdat.uni.sel', 'hyperdat.bi.sel', 
     'ppp.sps.uni', "ppp.sps.bi", 'trees', 'winlist', 'rmax',
      file='../data/data4peruanalysisv7.1.RData')
```

